---
title: "Studying healthy ranges"
output: html_notebook
---

Ken Lassesen generated a table of healthy ranges based on the 100+ samples people sent to his [cfsremission.com](https://cfsremission.com) site.


I'll set up the environment and then read data to generate two data frames: `healthy.kl` (with Ken's data) and `healthy.ps` with PSI data:

```{r}
library(tidyverse)

healthy.kl <- readxl::read_excel("../LassesenHealthyRanges.xlsx", na="NULL",col_types = c(rep("text",2),rep("numeric",10),rep("text",4),rep("numeric",4)))
head(healthy.kl)

healthy.ps <- readr::read_csv("../healthy_ranges.csv")

head(healthy.ps)
```

Hmm, looks like the max/min values are identical for many of the taxa.

Let's look at all the microbes in the `ps` list and see how they compare to the same one in the `kl` list.

```{r}

setdiff(healthy.kl$taxa, healthy.ps$taxa)  # note that the two lists include all the same taxa

healthy.kl %>% dplyr::select(taxa,maxcnt,mincnt,taxon,LowEstimate) %>% inner_join(select(healthy.ps,taxa,max,min), by = c("taxa","mincnt" = "min","maxcnt" ="max")) %>% knitr::kable(digits=2)
```

