---
title: "Studying healthy ranges"
output: html_notebook
---

Ken Lassesen generated a table of healthy ranges based on the 100+ samples people sent to his [cfsremission.com](https://cfsremission.com) site.

His healthy ranges calculations are based only on my data plus his reverse-engineered estimates of uBiome ranges. He used my skew calculation to guess at whether his ranges should be skewed more to the left or the right.


I'll set up the environment and then read data to generate two data frames: `healthy.kl` (with Ken's data) and `healthy.ps` with PSI data:

```{r}
library(tidyverse)

healthy.kl <- readxl::read_excel("../LassesenHealthyRanges.xlsx", na="NULL",col_types = c(rep("text",2),rep("numeric",10),rep("text",4),rep("numeric",4)))
head(healthy.kl)

healthy.ps <- readr::read_csv("../healthy_ranges.csv")

head(healthy.ps)
```

Hmm, looks like the max/min values are identical for many of the taxa.

Let's look at all the microbes in the `ps` list and see how they compare to the same one in the `kl` list.

```{r}

setdiff(healthy.kl$taxa, healthy.ps$taxa)  # note that the two lists include all the same taxa

healthy.kl %>% dplyr::select(taxa,maxcnt,mincnt,taxon,LowEstimate) %>% inner_join(select(healthy.ps,taxa,max,min), by = c("taxa","mincnt" = "min","maxcnt" ="max")) %>% knitr::kable(digits=2)
```


Here's how I calculate my own healthy ranges

```{r}
#write.csv(na.omit(mhg_range_for(subset_samples(people.norm,Site=="gut" & Condition == "Healthy"))),file="healthy_ranges.csv")

# returns a df with min and max values for all taxa in ps
# mhg_range_for <- function(ps) {
#   mx <- apply(otu_table(ps), 1, max)
#   mn <- apply(otu_table(ps), 1, min)
#   md <- apply(otu_table(ps), 1, median)
#   sd <- apply(otu_table(ps),1, sd)
#  # sk <- apply(otu_table(ps), 1, psych::skew)
#   av <- apply(otu_table(ps), 1, mean)
#   df <- data.frame(taxa = rownames(otu_table(ps)),max = mx, min = mn, median = md, mean = av, sd = sd)
#   df
#   
# }

# h <- mhg_range_for(subset_samples(people.norm,Site=="gut" & Condition == "Healthy"))
# 
# ggplot(data = h %>% arrange(desc(mean)) %>% top_n(20), aes(x=reorder(taxa,-max),y=mean)) + 
#   #geom_boxplot() +
#   geom_linerange( aes(ymin = min, ymax = max)) + 
#   geom_point(aes(y=median), color = "red") + 
#   geom_point(aes(y=mean), color = "blue") + 
#     theme(axis.text.x = element_text(size = 15, angle=90))  + 
#     scale_y_log10(labels=function(x)x/10000) +
#   labs(title="Healthy People", x="Top 20 Taxa", y = "Abundance (log10 %)")

```

